# Assignment 3 - Stroke Prediction

# Loading the dataset the data 
```{r}
data <- read.csv('data/healthcare-dataset-stroke-data.csv')
head(data)
```


# Data cleaning

## Checking the dimension of the data 
```{r}
dim(data) # 5110 rows X 12 columns
```


```{r}
# Checking if there is missing data in our data set
print("Structure of the data: ")
str(data) # Checking the structure of the dataframe
cat("\n")
```
```{r}
# After seeing the data structure, BMI was in character data type so we convert it into numeric (Double)
# Converting the data type of bmi from chr to num
data$bmi <- as.numeric(data$bmi)
cat("\n")
print("Summary of the dataframe: ")
summary(data)
```
We found that there are 201 NA for BMI

# Replacing NA values in BMI.


```{r}

boxplot(data$bmi, horizontal=TRUE,main="Boxplot of BMI",col="darkcyan")
hist(data$bmi,main="Distrbution of BMI", col="darkcyan", xlab = "BMI", ylab = "Count")

```
Replacing missing values with the median because the distribution is left skewed.
```{r}
# replacing the NA values in the dataset with the median
data$bmi <- ifelse(is.na(data$bmi), median(data$bmi, na.rm=TRUE), data$bmi)

# Checking the summary of the data again
print("Data Summary:")
summary(data)
cat("\n")

```
```{r}
cat("\n")
print("Number of N/A data in our data:")
sum(is.na(data)) # sum of all na in our dataset
```
# Handeling the outliers 
## Ploting the box plot to see the outliers.
```{r}
# Seeing the structure of our dataframe to find outliers
str(data)
```
## Box plot to visualize the outliers

```{r}
boxplot(data$age, horizontal=TRUE,main="Boxplot of Age",col="limegreen")
boxplot(data$avg_glucose_level, horizontal=TRUE,main="Boxplot of avg_glucose_level",col="darkmagenta")
boxplot(data$bmi, horizontal=TRUE,main="Boxplot of bmi",col="darkcyan")
```
# Histogram 
```{r}
hist(data$age,main="Distribution of Age",col="limegreen")
hist(data$avg_glucose_level,main="Distribution of avg_glucose_level",col="darkmagenta")
hist(data$bmi,main="Distribution of bmi",col="darkcyan")
```
Since our columns avg_glucose_level and bmi.

```{r}
variables_replace <- c("avg_glucose_level", "bmi")
colors_replace <- c("darkmagenta", "darkcyan")
# for loop to plot the box plots
for (i in 1:length(variables_replace)) {
  # Calculate the lower and upper bounds for outliers
  Q1 <- quantile(data[[variables_replace[i]]], 0.25)
  Q3 <- quantile(data[[variables_replace[i]]], 0.75)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  
  # Identify the outliers
  outliers <- which(data[[variables_replace[i]]] < lower_bound | data[[variables_replace[i]]] > upper_bound)
  
  # Replace outliers with the median value
  median_val <- median(data[[variables_replace[i]]])
  data[[variables_replace[i]]][outliers] <- median_val
  
  boxplot(data[[variables_replace[i]]], horizontal=TRUE, main=variables_replace[i], col=colors_replace[i], pch = 20)
  
}

```


# Data Visualization


# Barplot
## Gender
```{r}
# Bar plot for Gender
bar_color <- c("pink","blue", "purple")
barplot(table(data$gender), xlab = "Gender", ylab = "Count", main = "Distribution of Gender", col= bar_color)
```

## Married Status
```{r}
bar_color <- c("indianred","lightgreen")
barplot(table(data$ever_married), xlab = "Ever married", ylab = "Count", main = "Distribution of Maried Status", col= bar_color)
```

## Resident Type
```{r}
bar_color <- c("papayawhip","orchid")
barplot(table(data$Residence_type), xlab = "Resident Type", ylab = "Count", main = "Distribution of Resident Type", col= bar_color)
```

## Smoking status

```{r}
bar_color <- c("turquoise","wheat","tomato","thistle")
barplot(table(data$smoking_status), xlab = "Smoking Status", ylab = "Count", main = "Distribution of Smoking Status", col= bar_color)
```

# BOX PLOT

```{r}
# the variables and plot parameters being defined
variables <- c("age", "avg_glucose_level", "bmi")
xlabels <- c("Stroke", "Stroke", "Stroke")
ylabels <- c("Age", "Average Glucose Level", "BMI")
titles <- c("Distribution of Age by Stroke", "Distribution of Average Glucose Level by Stroke", "Distribution of BMI by Stroke")
legend_loc <- c("bottom","topleft", "bottom")
colors <- c("yellow", "orange")

# for loop to plot the box plots
for (i in 1:length(variables)) {
  boxplot(formula(paste(variables[i], "~ stroke")), data = data,
          xlab = xlabels[i], ylab = ylabels[i], main = titles[i],
          horizontal = FALSE, col = colors)
  
  legend(legend_loc[i], legend = c("Stroke", "No stroke"), fill = colors)
}
```


# HISTOGRAM 
```{r}
# Titles and labels for histograms
titles <- c("Distribution of Age", "Distribution of Average Glucose Level", "Distribution of BMI")
xlabels <- c("Age", "Average Glucose Level", "BMI")

# Colors for histograms
colors_histogram <- c(10, 3, 5)

# Loop through the variables
for (i in 1:length(variables)) {
  # Plot histogram
  hist(data[[variables[i]]], xlab = xlabels[i], main = titles[i], col = colors_histogram[i])
}
```


